---
title: "index"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## expstudies

***

There is an opportunity to build widely useful actuarial tools compatible with the mortality data format from VM-51 Appendix 4. The expstudies package provides a framework for working with policy data containing a unique policy key with start and end dates. Compute-intensive algorithms are implemented in Rcpp/C++ for high performance.

## Installation
The latest stable version can be installed from CRAN:
```{r eval = FALSE}
install.packages('expstudies')
```
The latest development version can be installed from github:
```{r eval = FALSE}
#install.packages("devtools")
devtools::install_github('ActuarialAnalyst/expstudies')
```

##expstudies Example
```{r include = FALSE}
library(pander)
```
```{r}
library(expstudies)
```
The expstudies package takes records with a start and end date - 
```{r, results = FALSE}
records
```
```{r, results = "asis", echo = FALSE}
pander::pandoc.table(records)
```
and transforms them into policy intervals.
```{r, results = FALSE}
addExposures(records)
```
```{r, results = "asis", echo = FALSE}
pander::pandoc.table(head(addExposures(records), 4))
```
Much more is possible with the package. Read about premium pattern, exact calendar years, and other topics in the articles tab. I would start by reading the article on "Creating Exposure Intervals".




