---
title: "Premium Pattern"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{premium_pattern}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r include = FALSE}
library(expstudies)
library(dplyr)
library(magrittr)
library(pander)
library(tidyr)
```

Suppose we have some monthly exposures that we would like to add premium data to.
```{r, results = "hide"}
exposures_PM <- addExposures(records, type = "PM")
head(exposures_PM)
```
```{r, results = "asis", echo = FALSE}
pander::pandoc.table(head(exposures_PM))
```
Simulated premium data "trans" comes with the package.
```{r, results = "hide"}
head(trans)
```
```{r, results = "asis", echo = FALSE}
pander::pandoc.table(head(trans))
```

The addStart function adds the start date of the appropriate exposure interval to the transactions. 
```{r, results = "hide"}
trans_with_interval <- addStart(trans, exposures_PM)
head(trans_with_interval)
```
```{r, results = "asis", echo = FALSE}
pander::pandoc.table(head(trans_with_interval))
```
We can group and aggregate by key and start_int to get unique transaction rows corresponding to intervals in exposures_PM. 
```{r, results = "hide"}
trans_to_join <- trans_with_interval %>% group_by(start_int, key) %>% summarise(premium = sum(amt))
head(trans_to_join)
```
```{r, results = "asis", echo = FALSE}
pander::pandoc.table(head(trans_to_join))
```
Then we can join this to the exposures using a left join without duplicating any exposures.
```{r, results = "hide"}
premium_study <- exposures_PM %>% left_join(trans_to_join, by = c("key", "start_int"))
head(premium_study, 10)
```
```{r, results = "asis", echo = FALSE}
pander::pandoc.table(head(premium_study, 10))
```
Change the NA values resulting from the join to zeros using an if_else.
```{r, results = "hide"}
premium_study <- premium_study %>% mutate(premium = if_else(is.na(premium), 0, premium))
head(premium_study, 10)
```
```{r, results = "asis", echo = FALSE}
pander::pandoc.table(head(premium_study, 10))
```
Now we are free to do any calculations we want. For a simple example we calculate the average premium in the first two policy months. Refer to the section on adding additional information for more creative policy splits. 
```{r, results = "hide"}
premium_study %>% filter(policy_month %in% c(1,2)) %>% group_by(policy_month) %>% summarise(avg_premium = mean(premium))
```
```{r, results = "asis", echo = FALSE}
pander::pandoc.table(premium_study %>% filter(policy_month %in% c(1,2)) %>% group_by(policy_month) %>% summarise(avg_premium = mean(premium)))
```
###Other Uses for addStart
Suppose we were interested in what the last premium paid by a policy was for some predictive analytics project. Again we left join the premium to the exposure frame.
```{r, results = "hide"}
previous_premium_unfilled <- exposures_PM %>% left_join(trans_to_join, by = c("key", "start_int"))
head(previous_premium_unfilled)
```
```{r, results = "asis", echo = FALSE}
pander::pandoc.table(head(previous_premium_unfilled))
```
This time we fill in NA values with the previous paid premium instead of 0. The first interval is NA because there are no prior premiums.
```{r, results = "hide"}
previous_premium <- previous_premium_unfilled %>%
tidyr::fill(premium, .direction = "down")
```
```{r, results = "asis", echo = FALSE}
pander::pandoc.table(head(previous_premium))
```
Data manipulations similar to this can be used to engineer features for anything varying with time: account values, guarantees, planned premiums, etc...
