---
title: "Mortality/Lapse Rate Calculations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{expstudies_intro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include = FALSE}
library(expstudies)
library(dplyr)
library(magrittr)
library(pander)

exposures <- addExposures(records)
head(exposures)

exposures_mod <- exposures %>% 
  inner_join(select(records, key, issue_age, gender), by = "key") %>%
  mutate(attained_age = issue_age + duration - 1)
```
Start with the exposures_mod frame from the "Creating Exposure Intervals" article.
```{r, results = "hide"}
head(exposures_mod)
```
```{r, results = "asis", echo = FALSE}
pander::pandoc.table(head(exposures_mod), split.table = Inf)
```

Add a death indicator and use a full exposure in the year of death when performing a mortality study.
```{r, results = "hide"}
exposures_mort <- exposures_mod %>% 
  group_by(key) %>% 
  mutate(exposure_mod = if_else(duration == max(duration), 1, exposure), 
         death_cnt = if_else(duration == max(duration), 1, 0)) 

tail(exposures_mort, 4)
```
```{r, results = "asis", echo = FALSE}
exposures_mort <- exposures_mort %>% ungroup() %>% select(-exposure, -issue_age)
pander::pandoc.table(tail(exposures_mort, 4), split.table = Inf)
```
We then aggregate by duration to calculate mortality rates.
```{r, results = "hide"}
duration_rate <- exposures_mort %>% 
  group_by(duration) %>% 
  summarise(q = sum(death_cnt)/sum(exposure_mod))

duration_rate
```
```{r, results = "asis", echo = FALSE}
pander::pandoc.table(duration_rate)
```
Or by attained age and gender.
```{r, results = "hide"}
attained_age_gender_rates <- exposures_mort %>% 
  group_by(attained_age, gender) %>% 
  summarise(q = sum(death_cnt)/sum(exposure_mod))

tail(attained_age_gender_rates)
```
```{r, results = "asis", echo = FALSE}
pander::pandoc.table(tail(attained_age_gender_rates))
```

##Expected mortality

Some mortality tables have been loaded in an easy to join format so that users can conduct A/E analysis.
```{r}
summary(expstudies::mortality_tables)
```
The "table" column includes the table identifier from the SOA website's collection of mortality tables.
```{r, results = "hide"}
head(mortality_tables$AM92$AM92_Ultimate)
```
```{r, results = "asis", echo = FALSE}
pander::pandoc.table(head(mortality_tables$AM92$AM92_Ultimate))
```
We join the mortality table to the data frame using the attained age.
```{r, results = "hide"}
head(left_join(exposures_mort, mortality_tables$AM92$AM92_Ultimate, by = "attained_age"))
```
```{r, results = "asis", echo = FALSE}
pander::pandoc.table(head(left_join(exposures_mort, select(mortality_tables$AM92$AM92_Ultimate, -table), by = "attained_age")), split.table = Inf)
```
